What paradigms can webcam eye-tracking be used for? Attempted replications of five ``classic'' cognitive science experiments

### 2025-03-20 manuscript.Rmd
Finishing going through the things I can address for R1

The third study was a partial replication attempt of the "looking group" condition of
@mannsVisualPairedcomparisonTask2000 (we did not replicate their priming condition). This experiment used the visual
paired-comparison, which involves presenting a previously-viewed image
and novel image together and measuring the proportion of time spent
looking at each image. The expected pattern of results is that
participants will look more at novel objects. They
@mannsVisualPairedcomparisonTask2000 hypothesized that this pattern of
behavior could be used to measure the strength of memories. If a viewer
has a weak memory of the old image, then they may look at the old and
new images roughly the same amount of time. They tested this in two
ways. First, they showed participants a set of images, waited five
minutes, and then paired those images with novel images. They found that
participants spent more time (58.8% of total time) looking at the novel
images. They then measured memory performance one day later and found
that participants were more likely to recall images that they had spent
less time looking at during the visual paired-comparison task the
previous day.

@@

Thus, most of the key findings replicated. What should we make of the finding that failed to replicate? Notably, the proportion of looks to the novel object on Day 1 (54% or 55%, depending on analysis) was numerically smaller than in the original (59%), potentially decreasing the ability to detect correlations with eye-gaze. If so, the question is whether that is due to lower-quality eye-gaze tracking. This seems unlikely, because calibration quality did not appear to substantially affect this result (or any of the others). Alternative explanation include the well-documented decline effect (replications tend to find smaller effect sizes than the original) and the common finding that subject accuracy is lower when not testing undergraduates at elite universities.

@@

Second, researchers need to take care with timing accuracy. Running studies on modern computers requires synchronizing multiple clocks, something that does not happen automatically. Failure to synchronize can result in substantial inaccuracies [@slim2022moving]. As indicated by the current work, the current jsPsych implementation of WebGazer.js has reasonably good timing synchronization. We urge researchers to avoid using the current release of WebGazer.js as is, and instead either use jsPsych, our patch to WebGazer.js (https://github.com/brownhci/WebGazer/commit/35301e56c192ef265bb270873a6526205f183711), or other software that has been specifically documented as addressing common issues in timing for web-based eye-tracking.
Third, it is true for any eyetracker that precise data depends on good calibration, and good calibration depends in part on lighting and background. We suspect that it is possible to give subjects automated feedback on how to improve their own calibration, which could result in higher-quality data and fewer subject exclusions based on poor calibration.

>>>>

The third study was a partial replication attempt of the "looking group" condition of
@mannsVisualPairedcomparisonTask2000 (we did not replicate their priming condition). This experiment used the visual
paired-comparison, which involves presenting a previously-viewed image
and novel image together and measuring the proportion of time spent
looking at each image. The expected pattern of results is that
participants will look more at novel objects. @mannsVisualPairedcomparisonTask2000 hypothesized that this pattern of
behavior could be used to measure the strength of memories. If a viewer
has a weak memory of the old image, then they may look at the old and
new images roughly the same amount of time. They tested this in two
ways. First, they showed participants a set of images, waited five
minutes, and then paired those images with novel images. They found that
participants spent more time (58.8% of total time) looking at the novel
images. They then measured memory performance one day later and found
that participants were more likely to recall images that they had spent
less time looking at during the visual paired-comparison task the
previous day.

@@

Thus, most of the key findings replicated. What should we make of the finding that failed to replicate? Notably, the proportion of looks to the novel object on Day 1 (54% or 55%, depending on analysis) was numerically smaller than in the original (59%), potentially decreasing the ability to detect correlations with eye-gaze. If so, the question is whether that is due to lower-quality eye-gaze tracking. This seems unlikely, because calibration quality did not appear to substantially affect this result (or any of the others). Alternative explanation include the well-documented decline effect [replications tend to find smaller effect sizes than the original, @schooler2011unpublished; @open2015estimating; @hartshorne2012tracking] and the common finding that subject accuracy is lower when not testing undergraduates at elite universities [@hartshorne2019meta; @germine2012web].

@@

Second, researchers need to take care with timing accuracy. Running studies on modern computers requires synchronizing multiple clocks, something that does not happen automatically. Failure to synchronize can result in substantial inaccuracies [@slim2022moving]. As indicated by the current work, the current jsPsych implementation of WebGazer.js has timing synchronization sufficient for many studies. We urge researchers to avoid using the current release of WebGazer.js as is, and instead either use jsPsych, our patch to WebGazer.js (https://github.com/brownhci/WebGazer/commit/35301e56c192ef265bb270873a6526205f183711), or other software that has been specifically documented as addressing common issues in timing for web-based eye-tracking.

Third, it is true for any eyetracker that precise data depends on good calibration, and good calibration depends in part on lighting and background [see, for instance, @yuksel2023investigation]. We suspect that it is possible to give subjects automated feedback on how to improve their own calibration, which could result in higher-quality data and fewer subject exclusions based on poor calibration.
